Fig2: Candidate genes

### 2020-05-03 analysis/figure-2.Rmd
new fig 2

```{r setup, message=F}
library(tidyverse)
library(ggtext)
library(cowplot)
library(ggpubr)
library(knitr)
library(kableExtra)
library(corrr)

@@

candidatevsdwide <- candidatevsd  %>%
    pivot_wider(names_from = gene, values_from = counts) 
FH <- subsetcandidatevsdwide("female", "hypothalamus")
FP <- subsetcandidatevsdwide("female", "pituitary")
FG <- subsetcandidatevsdwide("female", "gonad")
MH <- subsetcandidatevsdwide("male", "hypothalamus")
MP <- subsetcandidatevsdwide("male", "pituitary")
MG <- subsetcandidatevsdwide("male", "gonad") 
```

@@

  p <- candidatevsd %>%
    filter(tissue %in% whichtissue,
          gene %in% whichgenes,
          sex %in% whichsex) %>%
    mutate(treatment = factor(treatment, levels = charlevels)) %>%
    mutate(treatmentNum = as.numeric(treatment)) %>%
    ggplot(aes(x = treatmentNum, y = counts)) +
    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
    #geom_smooth(aes(color = sex)) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    facet_wrap(~sex, scales = "free_y", nrow = 2) +
    scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9),
                       labels = charlevels) +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme_B3() + 
    theme(legend.position = "none",
          axis.title.y = element_text(face = "italic"),
          axis.text.x = element_blank()) +
    labs(y = whichgenes,
         x = NULL) 
  return(p)
}

scattercorrelations <- function(df, gene1, gene2, mylinecolor, myxlab, myylab){
  p <- ggplot(df, aes(x = gene1, y = gene2)) +
    labs(subtitle = " ", x = myxlab, y = myylab) + 
    scale_color_manual(values = allcolors) +
    geom_smooth(method = "lm",  color = mylinecolor) +
    geom_point(aes(color = treatment))  +
    theme_B3() +  
    theme(axis.title = element_text(face = "italic"), 
            legend.position = "none") +
    stat_cor( size = 2)
  return(p)
}

@@

## plot sig candidate genes
a <- candidateboxplot("hypothalamus", "DRD1", sexlevels) + labs(x = NULL, title = "Hypothalamus")

g <- candidateboxplot("pituitary", "PRL", sexlevels) + labs(x = NULL, title = "Pituitary")

m <- candidateboxplot("gonad", "CREBRF", sexlevels) + labs(x = NULL, title = "Gonad") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

c <- plotcorrplot(hyp3, "females and males") + theme(legend.position = "none")  + labs(title = "")
i <- plotcorrplot(pit3, "females and males") + theme(legend.position = "none") + labs(title = "")
o <- plotcorrplot(gon3, "females and males") + theme(legend.position = "bottom")  + labs(title = "")

e <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female")+ labs(title = "")
k <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "female")+ labs(title = "")
q <- scattercorrelations(FG, FG$ESR2, FG$CRHBP, "#969696", "ESR2", "CRHBP")   + labs(subtitle = "female")+ labs(title = "")

f <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C") + theme(axis.title.y = element_blank()) + labs(subtitle = "male") + labs(title = "")
l <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1") + theme(axis.title.y = element_blank()) + labs(subtitle = "male") + labs(title = "")
r <- scattercorrelations(MG, MG$ESR2, MG$CRHBP, "#525252", "ESR2", "CRHBP")  + theme(axis.title.y = element_blank()) + labs(subtitle = "male") + labs(title = "")

plot_grid(a,c,e,f,g,i,k,l,m,o,q,r, nrow = 3, rel_widths = c(1,1.2,0.8,0.75), rel_heights = c(1,1,1.2),
          labels = "AUTO", label_size =  8)
```

>>>>

```{r setup, message=F}
library(tidyverse)
library(ggtext)
library(cowplot)
library(ggpubr)
library(knitr)
library(kableExtra)
library(corrr)
library(ggsignif)

@@

candidatevsdwide <- candidatevsd  %>%
    pivot_wider(names_from = gene, values_from = counts) 
FH <- subsetcandidatevsdwide("female", "hypothalamus")
FP <- subsetcandidatevsdwide("female", "pituitary")
FG <- subsetcandidatevsdwide("female", "gonad")
MH <- subsetcandidatevsdwide("male", "hypothalamus")
MP <- subsetcandidatevsdwide("male", "pituitary")
MG <- subsetcandidatevsdwide("male", "gonad")

@@

  p <- candidatevsd %>%
    filter(tissue %in% whichtissue,
          gene %in% whichgenes,
          sex %in% whichsex,
          treatment != "control") %>%
    mutate(treatment = factor(treatment, levels = charlevels)) %>%
    ggplot(aes(x = treatment, y = counts)) +
    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    facet_wrap(~gene, scales = "free_y", nrow = 2) +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme_B3() + 
    theme(legend.position = "none",
          strip.text.y = element_text(face = "italic"),
          strip.text.x = element_text(face = "italic"),
          axis.text.x = element_blank()) +
    labs(y = paste(whichsex, "gene expression", sep = " "),
         x = NULL) +
    geom_signif(comparisons = list(c( "bldg", "lay"),
                                 c( "lay", "inc.d3"),
                                 c("inc.d3", "inc.d9"),
                                 c( "inc.d9", "inc.d17"),
                                 c( "inc.d17", "hatch"),
                                 c("hatch", "n5"),
                                 c( "n5", "n9")),  
                map_signif_level=TRUE,
                textsize = 1.5, family = 'Helvetica',
                vjust = 1.5, size = 0)

  return(p)
}

scattercorrelations <- function(df, gene1, gene2, mylinecolor, myxlab, myylab){
  p <- ggplot(df, aes(x = gene1, y = gene2)) +
    labs(subtitle = " ", x = myxlab, y = myylab) + 
    scale_color_manual(values = allcolors) +
    geom_smooth(method = "glm",  color = mylinecolor) +
    geom_point(aes(color = treatment))  +
    theme_B3() +  
    theme(axis.title = element_text(face = "italic"), 
            legend.position = "none") +
    stat_cor( size = 2)
  return(p)
}
```

## plot sig candidate genes
a <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female") + labs(title = "Hypothalamus")
e <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "female")+ labs(title = "Pituitary")
i <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = "female")+ labs(title = "Gonad")

b <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male") + labs(title = "")
f <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = "male") + labs(title = "")
j <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = "male") + labs(title = "")

c <- candidateboxplot("hypothalamus", c("DRD1", "HTR2C"), "female") + labs(x = NULL, title = "") 
g <- candidateboxplot("pituitary", c( "AVPR1A", "CRHR1" ), "female") + labs(x = NULL, title = "")
k <- candidateboxplot("gonad", c("ESR1", "PGR"), "female") + labs(x = NULL, title = "") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

d <- candidateboxplot("hypothalamus",  c("DRD1", "HTR2C"), "male") + labs(x = NULL, title = " ") 
h <- candidateboxplot("pituitary", c("AVPR1A", "CRHR1"), "male") + labs(x = NULL, title = " ")
l <- candidateboxplot("gonad", c("ESR1", "PGR"), "male") + labs(x = NULL, title = " ") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

plot_grid(a,b,c,d,e,f,g,h,i,j,k,l, nrow = 3, 
          rel_widths = c(0.8,0.8,1,1), rel_heights = c(1,1,1.2),
          labels = c("A", " ", "B", "",
                     "C", " ", "D", "",
                     "E", " ", "F", ""), label_size =  8)

@@

## suplement correlations

```{r supplfig2, fig.width = 7.25, fig.height = 7.25}

a <- plotcorrplot(hyp1, "females") + theme(legend.position = "none", axis.text.x = element_blank())  + labs(title = "")
c <- plotcorrplot(pit1, "females") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(title = "")
e <- plotcorrplot(gon1, "females") + theme(legend.position = "bottom")  + labs(title = "")

b <- plotcorrplot(hyp2, "males") + theme(legend.position = "none", axis.text.x = element_blank())  + labs(title = "")
d <- plotcorrplot(pit2, "males") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(title = "")
f <- plotcorrplot(gon2, "males") + theme(legend.position = "bottom")  + labs(title = "")

plot_grid(a,b,c,d,e,f, nrow = 3, rel_heights = c(1,1,1.5))
